R package glmm

### 2014-06-02 CompareExamples/glmmexamples.tex
Added salamander results, but now thinking I need to fix something.

First, I fit the data using glmer in lme4. Then I fit the data using glmm with $m=10^4$. This took 16.75 minutes on my netbook. The point estimates and the standard errors were very similar between the two  methods of model-fitting. The documentation says that the default method of model-fitting uses the Laplace approximation, but Charlie wonders if they actually use numerical integration for simple problems like this.

First, the results from glmer:
\begin{verbatim}
> summary(gm1)
Generalized linear mixed model fit by maximum likelihood (Laplace
  Approximation) [glmerMod]
 Family: binomial ( logit )
Formula: cbind(incidence, size - incidence) ~ period + (1 | herd)
   Data: cbpp

@@

\section{Salamander}
I fit the salamander data set using  my glmm package and compared those results to the results of Yun Ju Sung and Charlie Geyer.

\begin{verbatim}
> set.seed(1234)
> sal.glmm3<-glmm(Mate~Cross,random=list(~0+Female,~0+Male),varcomps.names=c("F","M"),data=salamander,family.glmm=bernoulli.glmm,m=10^4)
> summary(sal.glmm3)

@@

>>>>

First, I fit the data using glmer in lme4. The model summary says it was fit with Adaptive Hermite Quadrature with 9 quadrature nodes. Then I fit the data using glmm with $m=10^4$. This took 16.75 minutes on my netbook. The point estimates and some of the standard errors are very similar between the two  methods of model-fitting.

\begin{tabular}{lccccc}
& Intercept & period2 & period3 & period4 & $\nu$ \\ \hline
glmer (numerical integration) & -1.40 & -.99 & -1.13 & -1.58 & .41\\
glmm $m=10^4$ & -1.42 & -.99 & -1.13 & -1.58 & .44 \\
\end{tabular}

Here are the full results from glmer:
\begin{verbatim}
> summary(gm1)
Generalized linear mixed model fit by maximum likelihood (Laplace
  Approximation) [glmerMod]
 Family: binomial ( logit )
Formula: cbind(incidence, size - incidence) ~ period + (1 | herd)
   Data: cbpp

@@

\section{Salamander}
The salamander dataset represents success or failure in matings between two types of salamanders denoted by R and W. Four types of crosses are possible: RR, RW, WR, WW. These are the fixed effects. There is a random effect for each female salamander and for each male salamander. Thus, there are two variance components: one for the females and one for the males. The response is binary: whether or not the mating was successful.

I fit the salamander data set using  my glmm package (with $m=10^4$ twice and with $m=10^6$ once)  and compared those results to the MCEM results of Booth and Hobert. They do not look so great, and the $m=10^6$ results took 41 hours to get.

\begin{tabular}{lcccccc}
& Intercept & RW & WR & WW & $\nu_F$ & $\nu_M$ \\ \hline
MCEM (B+H) & 1.03 & .32 & -1.95 & .99 & 1.4 & 1.25 \\
glmm $m=10^4$ & 1.11 & -.55 & -3.17& -.21 & 1.21 & .95 \\
glmm $m=10^4$ & .94 & -.76 & -2.72 & .0023 & 1.60 & .99 \\
glmm $m=10^6$ &  0.84&  -0.81&  -2.81&  -0.21&1.21& 1.02 \\
\end{tabular}

\begin{verbatim}
> set.seed(1234)
> sal.glmm3<-glmm(Mate~Cross,random=list(~0+Female,~0+Male),varcomps.names=c("F","M"),
data=salamander,family.glmm=bernoulli.glmm,m=10^4)
> summary(sal.glmm3)

@@

Here are the results with $m=10^6$. 
\begin{verbatim}
> ptm <- proc.time()
> sal.glmm3<-glmm(Mate~Cross,random=list(~0+Female,~0+Male),varcomps.names=c("F"
,"M"),data=salamander,family.glmm=bernoulli.glmm,m=10^6)
>  proc.time()-ptm
      user     system    elapsed 
148604.323    595.454 174153.241 
> summary(sal.glmm3)

Call:
glmm(fixed = Mate ~ Cross, random = list(~0 + Female, ~0 + Male), 
    varcomps.names = c("F", "M"), data = salamander, family.glmm = bernoulli.glm
m, 
    m = 10^6)

Fixed Effects:
            Estimate Std. Error z value Pr(>|z|)    
(Intercept)   0.8413     0.2756   3.052  0.00227 ** 
CrossRW      -0.8074     0.3737  -2.161  0.03072 *  
CrossWR      -2.8062     0.4097  -6.849 7.43e-12 ***
CrossWW      -0.2135     0.3887  -0.549  0.58289    
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Variance Components for Random Effects (P-values are one-tailed):
  Estimate Std. Error z value Pr(>|z|)/2    
F   1.2090     0.2443   4.948   3.74e-07 ***
M   1.0196     0.2013   5.065   2.04e-07 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

> sal.glmm3$beta.pql
[1]  0.83806013 -0.57917030 -2.44205490 -0.01778234
> sal.glmm3$nu.pql
[1] 1.122464 1.036371
